---
title: "Activity Prediction"
author: "Uma Srinivas Majji"
date: "10/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

## Loading Packages
```{r echo=TRUE, results='hide', message=FALSE,warning=FALSE}
library(caret)
library(dplyr)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
```

## Data Preparation

```{r}
# set the url for the download
urlTrain <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urlValid <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# download the datasets
training <- read.csv(url(urlTrain))
validation <- read.csv(url(urlValid))
training$classe <- as.factor(training$classe)

# create a partition with the training dataset
inTrain <- createDataPartition(y=training$classe,p=0.7,list=FALSE)
trainSet <- training[inTrain,]
testSet <- training[-inTrain,]
dim(trainSet)
dim(testSet)

# remove variables with near zero variance
NZV <- nearZeroVar(trainSet)
trainSet <- trainSet[,-NZV]
testSet <- testSet[,-NZV]

# remove variables that are mostly NA
allNA <- sapply(trainSet,function(x)mean(is.na(x)))>0.95
trainSet <- trainSet[,allNA==FALSE]
testSet <- testSet[,allNA==FALSE]

# remove identification only variables
trainSet <- trainSet[,-(1:5)]
testSet <- testSet[,-(1:5)]
dim(trainSet)
dim(testSet)
```

## Correlation Analysis

```{r}
#library(corrplot)
corMatrix <- cor(trainSet[,-54])
corrplot(corMatrix,order="FPC",method="color",type="lower",tl.cex=0.8,tl.col=rgb(0,0,0))
```


## Prediction Model building

### Method - Random Forest

```{r}
# model fit
set.seed(12345)
controlRF <- trainControl(method="cv",number=3,verboseIter=FALSE)
modFitRandForest <- train(classe~.,data=trainSet,method="rf",trControl=controlRF)
modFitRandForest$finalModel

# prediction on test set
predRandForest <- predict(modFitRandForest,testSet)
confMatRandForest <- confusionMatrix(predRandForest,testSet$classe)

# plot matrix results
plot(confMatRandForest$table,col=confMatRandForest$byClass,
     main=paste("Random Forest - Accuracy =",round(confMatRandForest$overall['Accuracy'],4)))
```

### Method - Decision Trees

```{r}
# model fit
set.seed(12345)
modFitDecTree <- rpart(classe~.,data=trainSet,method="class")
fancyRpartPlot(modFitDecTree)

# prediction on test data set
predDecTree <- predict(modFitDecTree,newdata=testSet,type="class")
confMatDecTree <- confusionMatrix(predDecTree,testSet$classe)
confMatDecTree

# plot matrix results
plot(confMatDecTree$table,col=confMatDecTree$byClass,
     main = paste("Decision Tree - Accuracy =",round(confMatDecTree$overall['Accuracy'], 4)))
```

### Method - Generalized Boosted Model

```{r}

# model fit
set.seed(12345)
controlGBM <- trainControl(method="repeatedcv",number=5,repeats=1)
modFitGBM <- train(classe~.,data=trainSet,method="gbm",trControl=controlGBM,verbose=F)
modFitGBM$finalModel

# prediction on test data set
predGBM <- predict(modFitGBM,newdata=testSet)
confMatGBM <- confusionMatrix(predGBM,testSet$classe)
confMatGBM

# plot matrix results
plot(confMatGBM$table,col=confMatGBM$byClass,
     main=paste("GBM - Accuracy =",round(confMatGBM$overall['Accuracy'],4)))
```

## Applying best fit model

```{r}
# Applying the Selected Model to the Validation Data
predValidation <- predict(modFitRandForest,validation)
predValidation
```